---
title: "Data Manipulation with `dplyr`"
output: github_document
---

Once you've imported data, you're going to beed to do some cleaning up.

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

Load the data and clean the names of variables.

```{r}
library(tidyverse)

options(tibble.print_min = 3)

litters_data = 
  read_csv("./data/FAS_litters.csv")
litters_data = 
  janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)
```


### `select`

Select specific variables.

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```

Select some variables in a range of columns.

```{r}
select(litters_data, group:gd_of_birth)
```

Delete some variables.

```{r}
select(litters_data, -pups_survive)
select(litters_data, -pups_survive, -group)
```

Select some variables and rename them.

```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

Rename some variables and show all the variables in a tibble.

```{r}
# litter_name=(overwrite the original data/base dataset) OR litter_name2=(old-fashioned)
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

Select variables that start or end with some words.

```{r}
select(litters_data, starts_with("gd"))
select(litters_data, ends_with("weight"))
```


Change the order of the variables.

```{r}
select(litters_data, litter_number, pups_survive, everything())
```


### `filter`

* `gd_of_birth == 20`
* `pups_born_alive >=2`
* `pups_survivie !=4`
* `!((pups_survivie == 4) & (gd_of_birth == 20))`
* `group %in% c("Con7", "Con8")` #useful for character variables
* `group == "Con7" & gd_of_birth == 20`

```{r}
filter(litters_data, gd_of_birth == 20)
```

```{r}
filter(litters_data, group == "Con7" & gd_of_birth == 20)
```

* `drop_na(litters_data)` will remove any row with a missing value.
* `drop_na(litters_data, wt_increase)` will remove rows for which `wt_increasing` is missing.
Use them carefully!!


### `mutate`

Create a new variable.

```{r}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       group = str_to_lower(group),
       # wt_gain_kg = wt_gain * 2.2
       )
```

### `arrange`

Sort data according to the values.

```{r}
head(arrange(litters_data, group, pups_born_alive), 10)
```

Sort in descending order.

```{r}
head(arrange(litters_data, desc(group), pups_born_alive), 10)
```

### `%>%` Piping

```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>% #shortcut: Ctrl+Shift+M
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(wt_gain)

# still run the code each line step by step

litters_data
```


Using piping with a non-tidyverse function (incomplete)

```{r}
litters_data %>% 
  lm(wt_gain ~ pups_born_alive, data = .) %>% 
  broom::tidy()
```



